\section{Data Points}

This section contains the list of claims that we want to make and the corresponding data points that would help support these claims.

\begin{enumerate}
\item {\bf Claim: Using models leads to compression.}
Data points: Compression ratio for at least one data set\\

\item {\bf Claim: Using models leads to faster query processing performance.}
Data points: Query evaluation time on base data vs model\\

\item {\bf Claim: Common learning algorithms are general enough to apply to common use cases without significant background knowledge.}
Data points: Set of parameters that have to be defined for a set of learning algorithms\\

\item {\bf Claim: One can realize the above benefits in compression and query processing performance through common learning algorithms (i.e., without designing a custom use-case specific model).}
Data points: Compression ratio and query evaluation time achieved by the same learning algorithm on at least two different datasets\\

\item {\bf Claim: The additive model representation improves query processing performance.}
Data points: Query evaluation time on additive model under varying confidence guarantees\\

\item {\bf Claim: Models improve data accuracy by removing noise.} 
Data points: Some metric comparing the noise of the measurements to that of the signal reconstructed from the model\\

\item {\bf Claim: Correlating signals can lead to even higher compression and query processing benefits.}
Data points: Compression ratio and query evaluation time achieved by using an individual model for each sensor vs using a model that correlates different sensors
\end{enumerate}